---
title: "Florida Retirement Age Demographics (2000–2020)"
---

```{r}
#| label: setup
#| message: false
#| echo: false

library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)

# Load county boundaries and decennial census age data
fl_geo <- read_rds("data/fl_geo.rds")
fl_age <- read_rds("data/fl_age_2000_2020.rds")

# Reshape data to wide format to compute percentage point changes (2000 to 2020)
age_wide <- fl_age |>
  mutate(pct = round(100 * (value / summary_value), 1)) |>
  select(GEOID, year, pct) |>
  pivot_wider(names_from = year, values_from = pct, names_prefix = "pct_") |>
  mutate(diff_age = round(pct_2020 - pct_2000, 1))

# Combine with county geometries and reproject to WGS84
fl_map <- left_join(fl_geo, age_wide, by = "GEOID") |>
  st_transform(4326)

# Create hover labels summarizing retirement-aged population for each county
labels <- sprintf(
  "<strong>%s</strong><br/>
   Ages 65+: %s%% (2000) &rarr; %s%% (2020)<br/>
   Change: %+0.1f percentage points",
  fl_map$NAME,
  fl_map$pct_2000, fl_map$pct_2020, fl_map$diff_age
) |> lapply(HTML)

# Define color scale for change in percentage of residents aged 65 and older
pal_age <- colorNumeric("YlOrRd", domain = fl_map$diff_age)
```

## Two Decades of Change in Florida's Retirement-Age Population

Between 2000 and 2020, Florida experienced a notable shift in the share of residents aged 65 and older (traditional retirement age):

- **Statewide Aging**: Many counties saw substantial increases in their retirement-aged population share over these two decades.
- **Geographic Variation**: Growth in the 65+ demographic was especially pronounced in central and coastal retirement destinations.

Hover over any county to view its 2000 share, 2020 share, and total percentage-point change.

```{r}
#| label: interactive-map
#| message: false
#| echo: false

leaflet(fl_map, width = "100%", height = "540px") |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal_age(diff_age),
    fillOpacity = 0.7, weight = 1, color = "white",
    label = labels,
    highlightOptions = highlightOptions(weight = 2, color = "#666", fillOpacity = 0.9, bringToFront = TRUE)
  ) |>
  addLegend(
    pal = pal_age,
    values = ~diff_age,
    title = "Change in Ages 65+<br/>(percentage points)",
    position = "bottomright"
  )
```
